Supplement to ``{\mytitle}''

### 2022-09-13 ms.Rnw
Changes to Model2 description, some responses to Ed, other minor edits

Individuals can progress to a latent infection $E_{u\vaccCounter}$ followed by symptomatic infection $I_{u\vaccCounter}$ with recovery to $R_{u\vaccCounter}$ or asymptomatic infection $A_{u\vaccCounter}$ with recovery to $R^A_{u\vaccCounter}$.
The force of infection depends on both direct transmission and an aquatic reservoir, $W_u(t)$, and is given by
\begin{equation}
\label{model2:lambda}
\lambda_{u}(t) = 0.5\big(1+\seasAmplitude \cos(2\pi t)\big)
\frac{\beta_W\, W_u(t)}{ \Wsat  + W_u(t)} +
\transmission \left\{\sum_{\vaccCounter=0}^4 I_{u\vaccCounter}(t) + \asymptomRelativeInfect \sum_{\vaccCounter=0}^4 A_{u\vaccCounter}(t) \right\}
\end{equation}
The latent state is therefore described by the vector $\bm{X}^{(2)}(t) = \big(S_{u\vaccCounter}(t),\allowbreak E_{u\vaccCounter}(t),\allowbreak I_{u\vaccCounter}(t),\allowbreak A_{u\vaccCounter}(t),\allowbreak R_{u\vaccCounter}(t),\allowbreak R_{u\vaccCounter}^A(t),\allowbreak W_u,\allowbreak u \in \seq{1}{U},\allowbreak \vaccCounter \in \seq{0}{4}\big)$.
The cosine term in Eq.~\myeqref{model2:lambda} accounts for annual seasonality, with a phase parameter $\phaseParm$.
The original implementation of Model~2 in \cite{lee20} fixes the phase at $\phaseParm = 0$.

@@

One approach to modeling a dynamic system is through probabilistic models.
With a probabilistic model, we suppose the existence of a joint density $f_{\bm{X}^{(\modelCounter)}_{\seq{0}{N}}, \bm{Y}^{(\modelCounter)}_{\seq{1}{N}}}$, with $\bm{X}^{(\modelCounter)}_{\seq{0}{N}}$ denoting the unobservable Markov process of model $\modelCounter$ at times $\seq{0}{N} = \{0, 1, \ldots, N\}$, and $\bm{Y}^{(\modelCounter)}_{\seq{1}{N}}$ denoting the observable process of the system at times $\seq{1}{N}$.
Under this framework, the observed data $y_{\seq{1}{n}}^*$ are assumed to be a single realization of the model $y_{\seq{1}{n}}^* \sim f_{\bm{X}^{(\modelCounter)}_{\seq{0}{N}}, \bm{Y}^{(\modelCounter)}_{\seq{1}{N}}}(x_{\seq{0}{N}}, y_{\seq{1}{N}}; \paramVec)$, where $\paramVec$ is a parameter vector that indexes the model.
Using a probabilistic model results in several advantages, including the ability to account for variability present in the system.
% which is of great interest to scientists and policy makers.
Furthermore, because each draw from the joint distribution represents a potential outcome of the dynamic system, best/worst case scenarios under the assumptions of the model can be easily obtained via simulation.

Once a model for the system has been proposed, the parameter vector $\paramVec$ needs to be estimated using the observed data.
There exist several algorithms, both frequentist and Bayesian, that can be used to obtain estimates of the parameters in stochastic dynamic models.
% including the EM algorithm, Kalman Filter (and extensions) \citep{evensen09}, and iterated filtering algorithms \citep{ionides15}.
In order to retain the ability to propose models that are scientifically meaningful rather than only those that are simply statistically convenient, we restrict ourselves to parameter estimation techniques that have the plug-and-play property, which is that the fitting procedure only requires the ability to simulate the latent process instead of evaluating transition densities \citep{breto09,he10}.
Plug-and-play algorithms include Bayesian approaches like ABC and PMCMC \citep{toni09,andrieu10}, but here we use frequentist methods to maximize model likelihoods.
To our knowledge, the only plug-and-play frequentist methods that can maximize the complete model likelihood are iterated filtering algorithms, which modify the well-known particle filter \citep{arulampalam02} by performing a random walk for each parameter and particle.
These perturbations are carried out iteratively over multiple filtering operations, using the collection of parameters from the previous filtering pass as the parameter initialization for the next iteration, and decreasing the random walk variance at each step.

@@

We perform a statistical test to determine whether or not the data indicate the presence of a linear trend in transmissibility.
To do this, we perform a profile-likelihood search on the parameter $\transmissionTrend$ and obtain a confidence interval via a Monte Carlo Adjusted Profile (MCAP) \citep{ionides17}.
Similar to Model~2, this model was originally implemented in two parts: an epidemic phase from October 2010 through March 2015, and an endemic phase from March 2015 onward.
As before, we continue to allow the re-estimation of some model parameters at the start of the endemic phase ($\reportRate, \sigmaProc^2$ and $\obsOverdispersion$) but require that the latent Markov process $X(t)$ carry over from one phase into the next.
The resulting confidence interval for $\transmissionTrend$ is $(\Sexpr{myround(mcap_results$ci[1], digits = 3)}, \Sexpr{myround(mcap_results$ci[2], digits = 3)})$, with the full results displayed in Fig.~\ref{fig:betat}.
These results are suggestive that the inclusion of a trend in transmission rate improves the quantitative ability of Model~1 to describe the observed data.
The reported results for Model~1 in the remainder of this article were obtained with the inclusion of the parameter $\transmissionTrend$.

@@

\citet{lee20} originally fit two versions of model~2 based on a presupposed change in cholera transmission from a epidemic phase to endemic phase that occurred in March, 2014.
We follow their decision to re-estimate model parameters at this break-point, but we include a requirement that the latent state $\bm{X}^{(2)}(t)$ at the start of the endemic period must be the same as the state at the end of the epidemic period.
This additional constraint is sensible for a mechanistic interpretation of the latent state, though it adds difficulty to the task of obtaining a model fit that closely resembles the observed data.
To combat this added difficulty, and to allow for a possible shift in the seasonality component in the force of infection, we introduce an additional phase parameter $\phaseParm$ in Eq.~\ref{model2:lambda}.
% \begin{equation}
% \lambda_{u}(t) = 0.5\big(1+\seasAmplitude \cos(2\pi t + \phaseParm)\big)
% \frac{\beta_W\, W_u(t)}{ \Wsat  + W_u(t)} +
% \beta \left\{\sum_{d=0}^4 I_{ud}(t) + \asymptomRelativeInfect \sum_{d=0}^4 A_{ud}(t) \right\}\label{eq:model2:phase}
% \end{equation}
% \eic{WE COULD SAVE AN EQUATION BY WRITING THE MORE GENERAL ONE INITIALLY AND SAYING THAT THEY ASSUMED $\phaseParm=0$}

% We also note that, without the inclusion of a phase parameter, there is an implicit assumption is that force of infection is greatest around January and weakest around June, and the addition of this parameter allows for the testing of this assumption.
We further increase model flexibility by fitting the parameter $\Wshed$ rather than treating it as fixed.
We implemented this model using the \code{spatPomp} \code{R} package \citep{asfaw21github}.
The model was then fit using the subplex algorithm, implemented in the  \code{subplex} package \citep{king2020Subplex}.
A comparison of the trajectory of the fitted model to the data is given in Fig~\ref{fig:mod2Traj}.

@@

\subsubsection{Model~3}
Model~3 is also of a probabilistic model.
In this model, both the latent and observable processes can be factored into department specific processes which interact with each other.
We denote the latent and measurement processes for Model~3 as $\bm{X}^{(3)}(t_{0:N}) = X^{(3)}_{1:U, 0:N}$, and $\bm{Y}^{(3)}(t_{1:N}) = Y_{1:U, 1:N}$ \eic{IS THIS NEEDED HERE?}.
The decision to model the system via metapopulation models versus a model aggregated to a larger spatial scale is one of great scientific interest, and evidence for the former approach has been provided in previous studies \citep{king15}.
Note that this evidence alone does not automatically discredit conclusions drawn via nationally aggregated models, as there are also good reasons to prefer a simple model over a complex one \citep{saltelli20,green15}.
Still, researchers that intend to use a mechanistic model to describe a dynamic system should design the model to be as realistic as their scientific understanding of the system and their computational abilities permit \eic{THIS SENTENCE SEEMS TO DISMISS THE ARGUMENTS IN THE PREVIOUS SENTENCE. PERHAPS THIS SENTENCE CAN JUST BE DELETED}.
Fitting scientifically flexible metapopulation models, however, remains a challenging statistical problem;
this is due to the fact that the approximation error of particle filters grows exponentially in the dimension of the model \citep{rebeschini15,park20}.
Algorithms that are based on the particle filter therefore become computationally intractable as the number of spatial units increase.

Parameters that must be fit in Model~3 are primarily shared between each department, the exception to this being the parameters $\beta_{W_u}$, and $\beta_u$, which are unique for each department $u \in \seq{1}{10}$.
To fit this model, \citet{lee20} simplified the parameter estimation problem by fitting independent department-level models to the data.
The shared parameters were calibrated using the cholera incidence data from Artibonite, and the department-specific parameters ($\beta_{W_u}$ and $\beta_u$) were fit using the data from their respective department.
Reducing a spatially coupled \eic{not spatially heterogeneous, as previously written} model to individual units in this fashion requires special treatment of any interactive mechanisms between spatial units, such as found in Eq.~\myeqref{eq:model3:foi}.
In particular, when considering a model for department $u$, the values $I_\nu(t)$ and $A_\nu(t)$, are unknown for $u \neq v \in \seq{1}{10}$.
A first order approximation of Eq.~\myeqref{eq:model3:foi} for each department $v \in \seq{1}{10}$ can be obtained using the weekly number of observed cholera cases in each department:
\begin{equation}
  I_v(t) + A_v(t) \approx \frac{365}{7\reportRate}y^*_v(t)\left(\frac{1}{\muDeath + \choleraDeath + \muIR} + \frac{1 - \symptomFrac}{\symptomFrac(\muDeath + \muIR)}\right)\label{eq:model3:approx}
\end{equation}
This approximation leads to department-specific models that are conditionally independent given the reported number of cholera infections in the remaining departments.
Here, we refer to a collection of POMP models that are independent across units as a PanelPOMP.

@@

# model 2 AIC
mod2_n_params <- 21  # TODO: adjust Number of parameters fit in Model 2
mod2_aic <- 2 * mod2_n_params - 2 * mod2_ll

@@

Obtaining forecasts in this fashion allows us to get results that are consistent with the model assumptions and the observed data.

@@

\jwc{PARAGRAPH REMOVED}
% Despite these results suggesting that cholera was likely to be eliminated from Haiti without an increase in vaccination effort, we in no way suggest that vaccinations shouldn't be used to combat cholera---or other infectious diseases---in other parts of the world.
% In fact, our analysis shows that an increase in vaccination does drastically reduce the negative effects of a worst-case scenario.
% Avoid comments on vaccination policy recommendations - Jesse

>>>>

Individuals can progress to a latent infection $E_{u\vaccCounter}$ followed by symptomatic infection $I_{u\vaccCounter}$ with recovery to $R_{u\vaccCounter}$ or asymptomatic infection $A_{u\vaccCounter}$ with recovery to $R^A_{u\vaccCounter}$.
The force of infection depends on both direct transmission and an aquatic reservoir, $W_u(t)$, and is given by
\begin{equation}
\label{model2:lambda}
\lambda_{u}(t) = 0.5\big(1+\seasAmplitude \cos(2\pi t + \phaseParm)\big)
\frac{\beta_W\, W_u(t)}{ \Wsat  + W_u(t)} +
\transmission \left\{\sum_{\vaccCounter=0}^4 I_{u\vaccCounter}(t) + \asymptomRelativeInfect \sum_{\vaccCounter=0}^4 A_{u\vaccCounter}(t) \right\}
\end{equation}
The latent state is therefore described by the vector $\bm{X}^{(2)}(t) = \big(S_{u\vaccCounter}(t),\allowbreak E_{u\vaccCounter}(t),\allowbreak I_{u\vaccCounter}(t),\allowbreak A_{u\vaccCounter}(t),\allowbreak R_{u\vaccCounter}(t),\allowbreak R_{u\vaccCounter}^A(t),\allowbreak W_u,\allowbreak u \in \seq{1}{U},\allowbreak \vaccCounter \in \seq{0}{4}\big)$.
The cosine term in Eq.~\myeqref{model2:lambda} accounts for annual seasonality, with a phase parameter $\phaseParm$.
The implementation of Model~2 in \cite{lee20} fixes $\phaseParm = 0$.

@@

One approach to modeling a dynamic system is through probabilistic models.
With a probabilistic model, we suppose the existence of a joint density $f_{\bm{X}^{(\modelCounter)}_{\seq{0}{N}}, \bm{Y}^{(\modelCounter)}_{\seq{1}{N}}}$, with $\bm{X}^{(\modelCounter)}_{\seq{0}{N}}$ denoting the unobservable Markov process of model $\modelCounter$ at times $\seq{0}{N} = \{0, 1, \ldots, N\}$, and $\bm{Y}^{(\modelCounter)}_{\seq{1}{N}}$ denoting the observable process of the system at times $\seq{1}{N}$.
Under this framework, the observed data $y_{\seq{1}{n}}^*$ are assumed to be a single realization of the model $y_{\seq{1}{n}}^* \sim f_{\bm{X}^{(\modelCounter)}_{\seq{0}{N}}, \bm{Y}^{(\modelCounter)}_{\seq{1}{N}}}(x_{\seq{0}{N}}, y_{\seq{1}{N}}; \paramVec)$, where $\paramVec$ is a parameter vector that indexes the model.
Using a probabilistic model results in several advantages, including the ability to account for variability present in the system.
% which is of great interest to scientists and policy makers.
Furthermore, because each draw from the joint distribution represents a potential outcome of the dynamic system, best/worst case scenarios under the assumptions of the model can be easily obtained via simulation.
The added benefits of a stochastic model are coupled with statistical difficulty of fitting the parameter vector $\paramVec$ to observed data.

There exist several algorithms, both frequentist and Bayesian, that can be used to obtain estimates of the parameters in stochastic dynamic models.
% including the EM algorithm, Kalman Filter (and extensions) \citep{evensen09}, and iterated filtering algorithms \citep{ionides15}.
In order to retain the ability to propose models that are scientifically meaningful rather than only those that are simply statistically convenient, we restrict ourselves to parameter estimation techniques that have the plug-and-play property, which is that the fitting procedure only requires the ability to simulate the latent process instead of evaluating transition densities \citep{breto09,he10}.
Plug-and-play algorithms include Bayesian approaches like ABC and PMCMC \citep{toni09,andrieu10}, but here we use frequentist methods to maximize model likelihoods.
To our knowledge, the only plug-and-play frequentist methods that can maximize the complete model likelihood are iterated filtering algorithms, which modify the well-known particle filter \citep{arulampalam02} by performing a random walk for each parameter and particle.
These perturbations are carried out iteratively over multiple filtering operations, using the collection of parameters from the previous filtering pass as the parameter initialization for the next iteration, and decreasing the random walk variance at each step.

@@

We perform a statistical test to determine whether or not the data indicate the presence of a linear trend in transmissibility.
To do this, we perform a profile-likelihood search on the parameter $\transmissionTrend$ and obtain a confidence interval via a Monte Carlo Adjusted Profile (MCAP) \citep{ionides17}.
Model~1 was implemented in \cite{lee20} by fitting two distinct phases: an epidemic phase from October 2010 through March 2015, and an endemic phase from March 2015 onward.
We similarly allow the re-estimation of process and measurement overdispersion parameters ($\sigmaProc^2$ and $\obsOverdispersion$), and require that the latent Markov process $X(t)$ carry over from one phase into the next.
The resulting confidence interval for $\transmissionTrend$ is $(\Sexpr{myround(mcap_results$ci[1], digits = 3)}, \Sexpr{myround(mcap_results$ci[2], digits = 3)})$, with the full results displayed in Fig.~\ref{fig:betat}.
These results are suggestive that the inclusion of a trend in transmission rate improves the quantitative ability of Model~1 to describe the observed data.
The reported results for Model~1 in the remainder of this article were obtained with the inclusion of the parameter $\transmissionTrend$.

@@

\cite{lee20} fit two versions of model~2 based on a presupposed change in cholera transmission from a epidemic phase to endemic phase that occurred in March, 2014.
The inclusion of a change-point in model states and parameters increases the flexibility of the model and hence the ability to fit the observed data.
The increase in model flexibility, however, results in hidden states that are inconsistent between model phases.
A naive solution to this would be to estimate only model parameters in separate phases and not the hidden states, but this may also result in hidden states in one phase of the model that are inconsistent with the other phase of the outbreak, since only the data from one phase is used to fit model parameters.

The inclusion of a model break-point in \cite{lee20} is perhaps due to a challenging feature of fitting a deterministic model via least squares: discrepancies between model trajectories and observed case counts in highly infectious periods of a disease outbreak will result in greater penalty than the discrepancies between model trajectories and observed case counts in times of relatively low infectiousness, resulting in a bias towards accurately describing periods of high infectiousness.
This issue is particularly troublesome for this case study: the inability to accurately fit times of low infectiousness may result in poor model forecasts, as few cases of cholera were observed the last few years epidemic.

To combat this issue, we fit the model to log-transformed case counts, where the difference between periods of high and low infectiousness is less drastic.
Other solutions include changing the measurement model to include overdispersion, as was done in Models~1 and~3.
Here we chose to fit the model to transformed case counts rather than adding overdispersion to the measurement model with the goal of making the fewest changes to the model in \cite{lee20} as possible.
In practice, however, the inclusion of overdispersion in the measurement process may be preferred \jwc{REFS???}.

We implemented this model using the \code{spatPomp} \code{R} package \citep{asfaw21github}.
The model was then fit using the subplex algorithm, implemented in the  \code{subplex} package \citep{king2020Subplex}.
A comparison of the trajectory of the fitted model to the data is given in Fig~\ref{fig:mod2Traj}.

@@

\subsubsection{Model~3}
Model~3 is also of a probabilistic model.
In this model, both the latent and observable processes can be factored into department specific processes which interact with each other.
% We denote the latent and measurement processes for Model~3 as $\bm{X}^{(3)}(t_{0:N}) = X^{(3)}_{1:U, 0:N}$, and $\bm{Y}^{(3)}(t_{1:N}) = Y_{1:U, 1:N}$ \eic{IS THIS NEEDED HERE?}.
The decision to model the system via metapopulation models versus a model aggregated to a larger spatial scale is one of great scientific interest, and evidence for the former approach has been provided in previous studies \citep{king15}.
Note that this evidence alone does not automatically discredit conclusions drawn via nationally aggregated models, as there are also good reasons to prefer a simple model over a complex one \citep{saltelli20,green15}.
% Still, researchers that intend to use a mechanistic model to describe a dynamic system should design the model to be as realistic as their scientific understanding of the system and their computational abilities permit \eic{THIS SENTENCE SEEMS TO DISMISS THE ARGUMENTS IN THE PREVIOUS SENTENCE. PERHAPS THIS SENTENCE CAN JUST BE DELETED}.

Fitting scientifically flexible metapopulation models is a challenging statistical problem.
In particular, parameter estimation techniques based on the particle filter become computationally intractable as the number of spatial units increase.
This is a result of the approximation error of particle filters growing exponentially in the dimension of the model \citep{rebeschini15,park20}.

Parameters that must be fit in Model~3 are primarily shared between each department, the exception to this being the parameters $\beta_{W_u}$, and $\beta_u$, which are unique for each department $u \in \seq{1}{10}$.
To avoid the parameter estimation issue in high-dimensional models, \citet{lee20} simplified the problem by fitting independent department-level models to the data.
The shared parameters were calibrated using the cholera incidence data from Artibonite, and the department-specific parameters ($\beta_{W_u}$ and $\beta_u$) were fit using the data from their respective department.
Reducing a spatially coupled model to individual units in this fashion requires special treatment of any interactive mechanisms between spatial units, such as found in Eq.~\myeqref{eq:model3:foi}.
In particular, when considering a model for department $u$, the values $I_\nu(t)$ and $A_\nu(t)$, are unknown for $u \neq v \in \seq{1}{10}$.
A first order approximation of Eq.~\myeqref{eq:model3:foi} for each department $v \in \seq{1}{10}$ can be obtained using the weekly number of observed cholera cases in each department:
\begin{equation}
  I_v(t) + A_v(t) \approx \frac{365}{7\reportRate}y^*_v(t)\left(\frac{1}{\muDeath + \choleraDeath + \muIR} + \frac{1 - \symptomFrac}{\symptomFrac(\muDeath + \muIR)}\right)\label{eq:model3:approx}
\end{equation}
This approximation leads to department-specific models that are conditionally independent given the reported number of cholera infections in the remaining departments.
Here, we refer to a collection of POMP models that are independent across units as a PanelPOMP.

@@

# model 2 AIC
mod2_n_params <- 16  # TODO: adjust Number of parameters fit in Model 2
mod2_aic <- 2 * mod2_n_params - 2 * mod2_ll

@@

Obtaining forecasts in this fashion allows us to obtain results that are consistent with the model assumptions and the observed data.

@@

### 2022-09-13 si/si.Rnw
Changes to Model2 description, some responses to Ed, other minor edits

In this model, the advantage afforded by vaccination is an increased probability that an infection is asymptomatic.
Therefore, under the assumptions of this model, all reported cases are assumed to be a fraction of individuals that transition from the exposed to the infected compartment, as noted in Eq.~\myeqref{model1:meas}.
\begin{equation}
  \label{model1:meas}
  y_{t} \mid \Delta N_{E_{k}I_{k}} = z_{t} \sim \text{NB}\left(\reportRate z_{t}, \obsOverdispersion \right)
\end{equation}

@@

Model~2 was fit via trajectory matching and therefore does not have an explicit measurement model.
Implicitly, however, this fitting process is equivalent to having a Gaussian measurement process.
Therefore the measurement model can be expressed as Eq.~\myeqref{model2:meas}.
\begin{equation}
  \label{model2:meas}
  y_{u, t} \mid \Delta N_{I_{ud}R_{ud}} = z_{u, t} \sim \text{N}\left(\reportRate z_{u, t}, \obsOverdispersion^2 \right)
\end{equation}

Where $\Delta N_{I_{ud}R_{ud}}$ is the number of individuals who move from compartment $I_{ud}$ to $R_{ud}$, for unit $u \in 1:U$ and vaccination cohort $d \in 0:5$.
We also note here that the variance parameter $\obsOverdispersion^2$ is separately fit for the epidemic and endemic phases.

@@

In this model, reported cholera cases are assumed to stem from individuals who develop symptoms and seek healthcare.
Therefore reported cases are assumed to come from an over-dispersed negative binomial model, given the increase in infected individuals:
\begin{equation}
  \label{model3:meas}
  y_{u, t} \mid \Delta N_{S_{ud}I_{ud}} = z_{u, t} \sim \text{NB}\left(\reportRate z_{u, t}, \obsOverdispersion \right)
\end{equation}

This measurement model is a minor change from that used in \cite{lee20}, which allowed for a change in the reporting rate on January 1st, 2018.
The fitted values of the reporting rate---before and after January 2018---were $0.97$ and $0.097$, respectively.
This major change in reporting rate alone could have been the cause that Model~3 originally failed to predict the eradication of cholera, as an overnight change from near perfect to almost non-existent reporting forces the model to explain the observed decrease in cases as a decrease in the reporting of cases rather than of prevalence of cases.
This shift was justified by a "change of the case definition that occurred on January 1st, 2018";
this claim was not cited, and we could find no evidence that such a drastic change in the reporting rate would be warranted.

>>>>

In this model, the advantage afforded by vaccination is an increased probability that an infection is asymptomatic.
Therefore, under the assumptions of this model, all reported cases are assumed to be a fraction of individuals that transition from the exposed to the infected compartment, as noted in Eq.~\myeqref{model1:meas}.
\begin{equation}
  \label{model1:meas}
  y_{t} \mid \Delta N_{E_{\vaccCounter}I_{\vaccCounter}}(t) = z(t) \sim \text{NB}\left(\reportRate z_(t), \obsOverdispersion \right)
\end{equation}

Where $y_t$ is the reported cholera cases at time $t \in 1:N$ and $\Delta N_{E_{\vaccCounter}I_{\vaccCounter}}(t)$ is the number of individuals who moved from compartment $E_\vaccCounter$ to $I_\vaccCounter$ since observation $t-1 \in 0:N-1$.

@@

As mentioned in the main text, Model~2 was fit using log-transformed observed cases.
We fit Model~2 using the subplex algorithm in the \code{subplex} package, using
a Gaussian measurement model (Eq.~\myeqref{model2:meas}) on the log-transformed cases as the loss function.

\begin{equation}
  \label{model2:meas}
  \log\big(y_{u, t} + 1\big) \mid \bm{X}^{(2)}(t) = \bm{x}^{(2)}(t) \sim \text{N}\left(\log \big(\reportRate \sum_{\vaccCounter} I_{u, \vaccCounter}(t) + 1\big), \obsOverdispersion^2 \right)
\end{equation}

Where $\sum_{\vaccCounter} I_{u, \vaccCounter}(t)$ is the total number of individuals who are infected at time $t$ in unit $u \in 1:U$, for all vaccination campaigns $\vaccCounter$.
We note that fitting a model with this loss function is equivalent to fitting using least squares, with $\log(y_{u, t} + 1)$ as the response variable.

@@

In this model, reported cholera cases are assumed to stem from individuals who develop symptoms and seek healthcare.
Therefore reported cases are assumed to come from an over-dispersed negative binomial model, given the increase in infected individuals:
\begin{equation}
  \label{model3:meas}
  y_{u, t} \mid \Delta N_{S_{u \vaccCounter}I_{u \vaccCounter}}(t) = z_{u}(t) \sim \text{NB}\left(\reportRate z_{u}(t), \obsOverdispersion \right)
\end{equation}

Where $\Delta N_{S_{u \vaccCounter}I_{u \vaccCounter}}(t)$ is the number of individuals who moved from compartment $S_{u \vaccCounter}$ to $I_{u \vaccCounter}$ since observation $t-1 \in 0:N-1$.

This measurement model is a minor change from that used in \cite{lee20}, which allowed for a change in the reporting rate on January 1st, 2018.
The fitted values of the reporting rate---before and after January 2018---were $0.97$ and $0.097$, respectively.
This major change in reporting rate alone could explain why Model~3 in \cite{lee20} failed to predict the eradication of cholera.
An overnight change from near perfect to almost non-existent reporting forces the model to explain the observed reduction in reported cases as a decrease in the reporting of cases, rather than a decrease in the prevalence of cholera.
This shift was justified by a "change of the case definition that occurred on January 1st, 2018";
this claim was not cited, and we could find no evidence that such a drastic change in the reporting rate would be warranted.
We therefore do not allow a change in reporting rate when fitting Model~3.
